An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2017-03-24 uai2017/paper.tex
Made minor changes, added reminders, now send to John

Figure~\ref{fig:gauss_mix_2} shows that {\sc MHminibatch} dominates in terms of
speed and efficiency. The histograms of the (final) minibatch sizes used each
iteration show that our method consumes significantly less data; the
distribution is short-tailed and the mean is 213, more than an order of
magnitude better compared to the other two methods (averages are 11962 and
63408). Sizes correspond to the running times of the methods, excluding the
likelihood computation of all data points for {\sc MHSubLhd}, which would
drastically increase running time.

@@

\subsection{SPARSE MATRIX FACTORIZATION}\label{ssec:collab_filtering}

@@

\begin{table}[t]
\caption{\textbf{TODO TABLE WITH SMF RESULTS}}
\small % Daniel: yeah, sorry, but otherwise it doesn't fit in a column
\label{tab:smf}
\vskip 0.15in
\begin{center}
\begin{tabular}{l l l l}
\textbf{RMSE} & \textbf{MovieLens 10m} & \textbf{Netflix}  \\
\hline \\
{\sc MHminibatch} & ??? & ??? & \\
ADAGrad           & ??? & ??? & \\
\end{tabular}
\end{center}
\vskip -0.1in
\end{table}

\textbf{Daniel: TODO this is our third experiment. Results are in progress...}

\textbf{We benchmark with the ADAGrad~\citep{adagrad} implementation of Sparse
Matrix Factorization described in~\citep{Canny:2013:BDA:2487575.2487677}, which
is a variant of Alternating Least Squares~\citep{Koren:2009:MFT:1608565.1608614}.
(Daniel: is this the correct citation for BIDMach? I assume this is what we will
use)}

\section{CONCLUSIONS}\label{sec:conclusion}

We have derived an M-H test for minibatch MCMC which approximates full data
tests. We present theoretical results and experimentally show the benefits of
our test on Gaussian mixtures, logistic regression, and sparse matrix
factorization.  Directions for future work include integrating our algorithm
with~\citep{cutting_mh_2014} by applying both tests each iteration or utilizing
the variance reduction techniques suggested in~\citep{DBLP:conf/icml/ChenG16}.
More elaborate extensions involve running more experiments on other models such
as neural networks and providing a recipe for how to use our algorithm
(following the framework of~\citep{sgmcmc_2015}).

\bibliography{paper}
\bibliographystyle{plainnat}

@@

This appendix is divided into two main parts. Appendix~\ref{app:proofs} provides
the proofs that we omitted from the main text due to space constraints.
Appendix~\ref{app:experiments} provides further details on the correction
distribution derivation and on our three main experiments to assist
understanding and reproducibility. Our code is open-source on
GitHub.\footnote{\url{https://github.com/BIDData/BIDMach/blob/master/src/main/scala/BIDMach/updaters/MHTest.scala}}
% Daniel: TODO remove for anonymous review later.

@@

\begin{table}[t]
\caption{Gaussian Mixture Model Statistics}
\small
\label{tab:poissons}
\vskip 0.15in
\begin{center}
\begin{tabular}{l l l l}
\textbf{Metric} & {\sc MHminibatch} & {\sc AustereMH} & {\sc MHSubLhd} \\
\hline
\T Equation~\ref{eq:log_prob_poissons} & -1430.0 & -1578.9 & -1232.7 \\
Chi-Squared & 3313.9 & 3647.7 & 2444.1 \\
\end{tabular}
\end{center}
\vskip -0.1in
\end{table}

@@

For large $n$, the per-bin distributions are approximated by Poissons with
parameter $\lambda_i=P_i n$. Given samples $\{\theta_1,\ldots,\theta_T\}$, let
$c_j$ denote the number of individual samples $\theta_i$ that fall in bin $j$
out of $N_{\rm bins}$ total. We have
\begin{equation}\label{eq:log_prob_poissons}
\log p(c_1, \ldots, c_{N_{\rm bins}} | P_1, \ldots, P_{N_{\rm bins}}) =
\sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)).
\end{equation}
Table~\ref{tab:poissons} shows the likelihoods. To facilitate interpretation we
perform significance tests using Chi-Squared distribution (also in
Table~\ref{tab:poissons}). Scores lie between~\citep{cutting_mh_2014}
and~\citep{icml2014c1_bardenet14}, but the variance of these values is high and
ordering changes depending on the range of samples generated. \textbf{Daniel:
one reviewer had a problem with this, and I agree. The easiest fix here is to
simply report standard deviation values. Also, we should emphasize that we are
not aiming to get the best score by this metric but simply something that
exhibits some form of ``correctness.''}

@@

\subsection{SPARSE MATRIX FACTORIZATION EXPERIMENT (SECTION~\ref{ssec:collab_filtering})}\label{app:collab_filtering}

\textbf{Daniel: TODO}

>>>>

Figure~\ref{fig:gauss_mix_2} shows that {\sc MHminibatch} dominates in terms of
speed and efficiency. The histograms of the (final) minibatch sizes used each
iteration show that our method consumes significantly less data; the
distribution is short-tailed and the mean is 172, more than an order of
magnitude better compared to the other two methods (averages are 12562 and
675008). \textbf{Daniel: Xinlei, these numbers don't match
Figure~\ref{fig:gauss_mix_2} and the figures need to be updated so the scatter
plot has the correct method names and the scaling key, while the histograms have
the correct names and the seaborn background.} We
further ran 10 runs of mixture of Gaussians experiments, and the average and
standard deviation values of the mean minibatch size for {\sc MHminibatch}, {\sc
AustereMH(c)}, {\sc MHSubLhd} are listed in table ~\ref{tab:gaussianres}. Sizes
correspond to the running times of the methods, excluding the likelihood
computation of all data points for {\sc MHSubLhd}, which would drastically
increase running time.

\begin{table}[t]
\caption{Average minibatch sizes ($\pm$ one standard deviation) on the Gaussian
mixture model. The averages are taken over 10 independent trials (3000 samples
each). \textbf{Daniel: Xinlei, can you add in one more decimal point for all the
numbers in this table, like I do in the Logistic Regression table?}}
% Daniel: this `small` command was carried over from the LR table which couldn't
% fit in a column. It's not needed here.
%\small 
\label{tab:gaussianres}
\vskip 0.15in
\begin{center}
\begin{tabular}{l l l l}
\textbf{Method} & \textbf{Average of MB Sizes} \\
\hline \\
{\sc MHminibatch}   & $182\pm 12$ \\
{\sc AustereMH(c)}  & $13540\pm 1521$ \\
{\sc MHSubLhd}      & $65759\pm 3222$ \\
\end{tabular}
\end{center}
\vskip -0.1in
\end{table}

@@

@@

\section{CONCLUSIONS AND DISCUSSIONS}\label{sec:conclusion}

We have derived an M-H test for minibatch MCMC which approximates full data
tests. We present theoretical results and experimentally show the benefits of
our test on Gaussian mixtures and a variety of logistic regression experiments.

The most pressing direction for future work is to integrate our M-H test with
Hamiltonian Monte Carlo and Langevin Dynamics methods to enable MCMC methods to
scale to models such as deep neural networks. With these types of models, our
proposal probabilities will almost certainly have to be asymmetric, but these
typically result in an estimate plus Gaussian noise which we can absorb into our
test.

Other possibilities for future work include integrating our algorithm
with~\citep{cutting_mh_2014} by applying both tests each iteration, utilizing
the variance reduction techniques suggested in~\citep{DBLP:conf/icml/ChenG16},
and providing recipe for how to use our algorithm following the framework
of~\citep{sgmcmc_2015}.

\newpage
\bibliography{paper}
\bibliographystyle{plainnat}

@@

This appendix is divided into two main parts. Appendix~\ref{app:proofs} provides
the proofs that we omitted from the main text due to space constraints.
Appendix~\ref{app:experiments} provides further details on the correction
distribution derivation and on our three main experiments to assist
understanding and reproducibility. 
% Daniel: TODO can decide whether to include this here or not. It might actually
% be better to do this in the logistic regression section.
% Our code is open-source on
% GitHub.\footnote{\url{https://github.com/BIDData/BIDMach/blob/master/src/main/scala/BIDMach/updaters/MHTest.scala}}

@@

\begin{table}[t]
\caption{Gaussian Mixture Model Statistics \textbf{TODO Xinlei, please fix this
table.}}
\small
\label{tab:poissons}
\vskip 0.15in
\begin{center}
\begin{tabular}{l l l l}
\textbf{Metric} & {\sc MHminibatch} & {\sc AustereMH} & {\sc MHSubLhd} \\
\hline
\T Equation~\ref{eq:log_prob_poissons} & -1430.0 & -1578.9 & -1232.7 \\
Chi-Squared & 3313.9 & 3647.7 & 2444.1 \\
\end{tabular}
\end{center}
\vskip -0.1in
\end{table}

@@

For large $n$, the per-bin distributions are approximated by Poissons with
parameter $\lambda_i=P_i n$. Given samples $\{\theta_1,\ldots,\theta_T\}$, let
$c_j$ denote the number of individual samples $\theta_i$ that fall in bin $j$
out of $N_{\rm bins}$ total. We have
\begin{equation}\label{eq:log_prob_poissons}
\log p(c_1, \ldots, c_{N_{\rm bins}} | P_1, \ldots, P_{N_{\rm bins}}) =
\sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)).
\end{equation}
Table~\ref{tab:poissons} shows the likelihoods. To facilitate interpretation we
perform significance tests using Chi-Squared distribution (also in
Table~\ref{tab:poissons}). Scores lie between~\citep{cutting_mh_2014}
and~\citep{icml2014c1_bardenet14}, but the variance of these values is high and
ordering changes depending on the range of samples generated. \textbf{Daniel:
Xinlei, one reviewer had a problem with this, and I agree. The easiest fix here
is to simply report standard deviation values. Also, we should emphasize that we
are not aiming to get the best score by this metric but simply something that
exhibits some form of ``correctness.''}

@@
